---
title: "a3_task4_laurel_wee"
author: "Laurel Wee"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(kableExtra)
library(lubridate)
library(janitor)
```
Assignment 3 task 4

# Summary

### Read in urchins data
```{r}
purple_urchin_naples <- read.csv(here("data","urchins.csv")) %>% 
  clean_names() %>% 
  mutate(naples_date = mdy(date)) %>% 
  mutate(naples_year= year(naples_date)) %>% 
  mutate(naples_month= month(naples_date)) %>% 
  filter(common_name == "Purple Urchin",  naples_year== "2010", naples_month== "1", site== "NAPL")
```

### Make and consider histogram and QQ-plots for the size distribution for the treatments "CONTROL" and "Annual". 

```{r}
# histogram for size distribution in each treatment
ggplot(purple_urchin_naples, aes(size, fill= treatment, bin= 30))+
  geom_histogram(color= "black")+
  scale_fill_manual(values = c("brown", "green"))+
  facet_wrap(~treatment)+
  labs(x= "Urchin Diameter (cm)", y= "Number of Urchins", title = "Size comparisons of Purple and Red Urchins in areas with or without giant kelp", subtitle = "CONTROL (no removal) vs ANNUAL (annual removal), Naples Reef, January 2010 (WEE)")+
  theme(legend.position = "none")

# qq-plot for size distribution in each treatment
ggplot(purple_urchin_naples, aes(sample= size, color= treatment))+
  geom_qq()+
  facet_wrap(~treatment)+
  scale_color_manual(values = c("brown", "green"))+
  labs(y="Urchin Diameter (cm)" , title = "Distributions of Purple and Red Urchins size in areas with or without giant kelp", subtitle = "CONTROL (no removal) vs ANNUAL (annual removal), Naples Reef, January 2010 (WEE)")+
  theme(legend.position = "none")
  
```

### Main takeaways about the distribution of urcin sizes for "CONTROL" and "ANNUAL"

## c. Create a table that includes summary statistics for both treatments of Purple urchins in Naples Reef in January 2010.
Table includes mean size (cm), median size (cm), standard deviation of the size (cm), and the sample size. 

```{r}
purple_urchin_summary_naples <- purple_urchin_naples %>%
  group_by(treatment) %>% 
  summarize(mean = mean(size),
            medium = median(size),
            sd = sd(size),
            n = n()) 

View(purple_urchin_summary_naples)

```

## d. Two sample t-test


```{r}
control_sample_urchins <- 
```

